---
title: "2_Gtex_DTU"
author: "Jeroen Gilis"
date: "02/11/2020"
output: html_document
---

**In order to run this script (2_Gtex_DTU.Rmd), the dataset**
**GTEx_benchmark_datasets_count.Rdata is required.** This file can be downloaded 
from Zenodo.  

Run the DTU analyses for all 5 methods on the Gtex benchmark dataset. Note that 
for DEXSeq and DRIMSeq we do not run the datasets with 50 samples in each group, 
as these methods do not scale to large datasets.

**If you do not want to run this script, its output can also be downloaded** 
**from Zenodo: GTEx_DTU_results_count.Rdata **

```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Load DTU methods

Source file that allows running all seven DTU methods that were assessed in this paper: satuRn, DoubleExpSeq, limma diffsplice, edgeRdiffsplice, DEXSeq, DRIMseq and NBSplice.

```{r, message=FALSE, warning=FALSE}
### load libraries
library(doMC)
library(edgeR)
library(limma)
library(DEXSeq)
#devtools::install_github("statOmics/satuRn", force = TRUE, quiet = FALSE)
# version 1.7.3 
library(satuRn)

source(file="./Performance_benchmarks/DTU_methods.R")
```

# Set up parallel execution

```{r}
nrCores <- 1

if(nrCores != 1) {
    doParallel <- TRUE
    doProgress <- 'none'
    registerDoMC(cores = nrCores)
} else {
    doParallel <- FALSE
    doProgress <- 'text'
}
```

# Run all analyses 

```{r}
### Load the benchmark datasets
load(file="./Data/GTEx_benchmark_datasets_count.Rdata")

### Run DTU analyses on benchmark data
print("start satuRn")
tStart <- Sys.time()
suppressWarnings(gtexDtuBenchmark_satuRn <- plyr::llply(
    .data = gtexBenchmarkDataLenient,
    .parallel = doParallel,
    .progress = doProgress,
    .inform = TRUE,
    .fun = function(localData) {
        ### Perform DTU analysis
        localRes <- satuRn_DTU(
            countData  = localData$data,
            tx2gene    = localData$metaInfo,
            sampleData = localData$design
        )

        ### Massage
        localRes$gene_modified <- localData$metaInfo$gene_modified[match(
            localRes$TXNAME, localData$metaInfo$TXNAME)]

        ### Return result
        return(list(dtuAnalysis = localRes))
    }
))
difftime(Sys.time(), tStart)
    
print("start edgeR_diffsplice")
tStart <- Sys.time()
gtexDtuBenchmark_edgeRdiffsplice <- plyr::llply(
    .data = gtexBenchmarkDataLenient,
    .parallel = doParallel,
    .progress = doProgress,
    .inform = TRUE,
    .fun = function(localData) {
        ### Perform DTU analysis
        localRes <- edgeR_diffsplice_DTU(
            countData  = localData$data,
            tx2gene    = localData$metaInfo,
            sampleData = localData$design
        )

        ### Massage
        localRes$gene_modified <- localData$metaInfo$gene_modified[match(
            localRes$TXNAME, localData$metaInfo$TXNAME)]

        ### Return result
        return(list(dtuAnalysis = localRes))
    }
)
difftime(Sys.time(), tStart)

print("start edgeR_offset")
tStart <- Sys.time()
gtexDtuBenchmark_edgeR_offset <- plyr::llply(
    .data = gtexBenchmarkDataLenient,
    .parallel = doParallel,
    .progress = doProgress,
    .inform = TRUE,
    .fun = function(localData) {
        ### Perform DTU analysis
        localRes <- edgeR_offset_DTU(
            countData  = localData$data,
            tx2gene    = localData$metaInfo,
            sampleData = localData$design)
        ### Massage
        localRes$gene_modified <- localData$metaInfo$gene_modified[match(
            localRes$TXNAME, localData$metaInfo$TXNAME)]
        ### Return result
        return(list(dtuAnalysis = localRes))
    })
difftime(Sys.time(), tStart)

print("start DESeq2_offset")
tStart <- Sys.time()
gtexDtuBenchmark_DESeq2_offset <- plyr::llply(
    .data = gtexBenchmarkDataLenient,
    .parallel = doParallel,
    .progress = doProgress,
    .inform = TRUE,
    .fun = function(localData) {
        ### Perform DTU analysis
        localRes <- DESeq2_offset_DTU(
            countData  = localData$data,
            tx2gene    = localData$metaInfo,
            sampleData = localData$design)
        ### Massage
        localRes$gene_modified <- localData$metaInfo$gene_modified[match(
            localRes$TXNAME, localData$metaInfo$TXNAME)]
        ### Return result
        return(list(dtuAnalysis = localRes))
    })
difftime(Sys.time(), tStart)
 
print("start DEXSeq")
tStart <- Sys.time()
gtexDtuBenchmark_DEXSeq <- plyr::llply(
    .data = gtexBenchmarkDataLenient,
    .parallel = doParallel,
    .progress = doProgress,
    .inform = TRUE,
    .fun = function(localData) {
            
        if(ncol(localData$data) > 40) {
            return(NULL) 
        } # do not run DEXSeq for datasets with many samples - too slow
          
        ### Perform DTU analysis
        localRes <- DEXSeq_DTU(
            countData  = localData$data,
            tx2gene    = localData$metaInfo,
            sampleData = localData$design
        )

        ### Massage
        localRes$gene_modified <- localData$metaInfo$gene_modified[match(
            localRes$TXNAME, localData$metaInfo$TXNAME)]

        ### Return result
        return(list(dtuAnalysis = localRes))
    }
)
difftime(Sys.time(), tStart)

### add method name to list names for easy post-analysis
names(gtexDtuBenchmark_satuRn) <- paste0('satuRn_',names(gtexDtuBenchmark_satuRn))
names(gtexDtuBenchmark_edgeRdiffsplice) <- paste0('edgeR_diffsplice_', names(gtexDtuBenchmark_edgeRdiffsplice))
names(gtexDtuBenchmark_DEXSeq) <- paste0('DEXSeq_', names(gtexDtuBenchmark_DEXSeq))
names(gtexDtuBenchmark_edgeR_offset) <- paste0('edgeR_offset_', names(gtexDtuBenchmark_edgeR_offset))
names(gtexDtuBenchmark_DESeq2_offset) <- paste0('DESeq2_offset_', names(gtexDtuBenchmark_DESeq2_offset))

### Save result
save(gtexDtuBenchmark_satuRn,
     gtexDtuBenchmark_edgeRdiffsplice,
     gtexDtuBenchmark_DEXSeq,
     gtexDtuBenchmark_edgeR_offset,
     gtexDtuBenchmark_DESeq2_offset,
     file="./Data/GTEx_DTU_results_count_new.Rdata")
```

```{r}
load("./Data/GTEx_DTU_results_count_new.Rdata")
```

