---
title: "2_Hsapiens_DTU"
author: "Jeroen Gilis"
date: "06/11/2020"
output: html_document
---

**In order to run this script (Hsapiens_DTU.Rmd), the dataset**
**Hsapiens_benchmark_datasets_count.Rdata  is required.** This file can be 
downloaded from Zenodo.

Here we run the DTU analyses for 5 DTU methods on the Hsapiens benchmark 
dataset.

**If you do not want to run this script, its output can also be downloaded**
**from Zenodo: Hsapiens_DTU_results_count.Rdata**

```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Load DTU methods

Source file that allows running all seven DTU methods that were assessed in 
this paper: satuRn, DoubleExpSeq, limma diffsplice, edgeRdiffsplice, DEXSeq, 
DRIMseq and NBSplice.

```{r, message=FALSE, warning=FALSE}
### load libraries
library(edgeR)
library(DEXSeq)
library(satuRn)
library(doMC)

source(file="./Performance_benchmarks/DTU_methods.R")
```

# Set up parallel execution

Note that the optimal parameter setting for parallelizing the DTU analyses depends on your specific computing system.

```{r}
nrCores <- 1

if(nrCores != 1) {
    doParallel <- TRUE
    doProgress <- 'none'

    registerDoMC(cores = nrCores)
} else {
    doParallel <- FALSE
    doProgress <- 'text'
}
```

# Run the analysis for all methods on both Hsapiens datasets

Run the DTU analyses for 7 DTU methods on both Hsapiens benchmark datasets.
This code runs approximately 30 minutes on a MacBook Pro 2018, processor; 
2,3 GHz Quad-Core Intel Core i5, 16GB RAM.

```{r}
### Load benchmark data
load(file="./Data/Hsapiens_benchmark_datasets_count.Rdata")
HsBenchmarkLenient <- list(HsBenchmarkLenient)
names(HsBenchmarkLenient)   <- paste0(names(HsBenchmarkLenient)  , 'filterLenient')

### Run DTU analysis on benchmark data
print("start satuRn")
tStart <- Sys.time()
suppressWarnings(HsDtuBenchmark_satuRn <- plyr::llply(
    .data = HsBenchmarkLenient,
    .inform = TRUE,
    .fun = function(localData) {
        ### Perform DTU analysis
        localRes <- satuRn_DTU(
            countData  = localData$data,
            tx2gene    = localData$metaInfo,
            sampleData = localData$design)
        ### Massage
        localRes$gene_modified <- localData$metaInfo$gene_modified[match(
            localRes$TXNAME, localData$metaInfo$TXNAME)]
        ### Return result
        return(list(dtuAnalysis = localRes))
    }
))
difftime(Sys.time(), tStart)

print("start edgeR_diffsplice")
tStart <- Sys.time()
HsDtuBenchmark_edgeRdiffsplice <- plyr::llply(
    .data = HsBenchmarkLenient,
    .parallel = doParallel,
    .progress = doProgress,
    .inform = TRUE,
    .fun = function(localData) {
        ### Perform DTU analysis
        localRes <- edgeR_diffsplice_DTU(
            countData  = localData$data,
            tx2gene    = localData$metaInfo,
            sampleData = localData$design)
        ### Massage
        localRes$gene_modified <- localData$metaInfo$gene_modified[match(
            localRes$TXNAME, localData$metaInfo$TXNAME)]
        ### Return result
        return(list(dtuAnalysis = localRes))
    }
)
difftime(Sys.time(), tStart)

print("start edgeR_offset")
tStart <- Sys.time()
HsDtuBenchmark_edgeR_offset <- plyr::llply(
    .data = HsBenchmarkLenient,
    .parallel = doParallel,
    .progress = doProgress,
    .inform = TRUE,
    .fun = function(localData) {
        ### Perform DTU analysis
        localRes <- edgeR_offset_DTU(
            countData  = localData$data,
            tx2gene    = localData$metaInfo,
            sampleData = localData$design)
        ### Massage
        localRes$gene_modified <- localData$metaInfo$gene_modified[match(
            localRes$TXNAME, localData$metaInfo$TXNAME)]
        ### Return result
        return(list(dtuAnalysis = localRes))
    })
difftime(Sys.time(), tStart)


print("start DESeq2_offset")
tStart <- Sys.time()

HsDtuBenchmark_DESeq2_offset <- plyr::llply(
    .data = HsBenchmarkLenient,
    .parallel = doParallel,
    .progress = doProgress,
    .inform = TRUE,
    .fun = function(localData) {
        ### Perform DTU analysis
        localRes <- DESeq2_offset_DTU(
            countData  = localData$data,
            tx2gene    = localData$metaInfo,
            sampleData = localData$design)
        ### Massage
        localRes$gene_modified <- localData$metaInfo$gene_modified[match(
            localRes$TXNAME, localData$metaInfo$TXNAME)]
        ### Return result
        return(list(dtuAnalysis = localRes))
    })
difftime(Sys.time(), tStart)


print("start DEXSeq")
tStart <- Sys.time()
HsDtuBenchmark_DEXSeq <- plyr::llply(
    .data = HsBenchmarkLenient,
    .parallel = doParallel,
    .progress = doProgress,
    .inform = TRUE,
    .fun = function(localData) {
        ### Perform DTU analysis
        localRes <- DEXSeq_DTU(
            countData  = localData$data,
            tx2gene    = localData$metaInfo,
            sampleData = localData$design)
        ### Massage
        localRes$gene_modified <- localData$metaInfo$gene_modified[match(
            localRes$TXNAME, localData$metaInfo$TXNAME)]
        ### Return result
        return(list(dtuAnalysis = localRes))
    }
)
difftime(Sys.time(), tStart)

### add method name to list names for easy post-analysis
names(HsDtuBenchmark_satuRn) <- paste0('satuRn_', names(HsDtuBenchmark_satuRn))
names(HsDtuBenchmark_DEXSeq)  <- paste0('DEXSeq_', names(HsDtuBenchmark_DEXSeq))
names(HsDtuBenchmark_edgeRdiffsplice) <- paste0('edgeR_diffsplice_', names(HsDtuBenchmark_edgeRdiffsplice))
names(HsDtuBenchmark_edgeR_offset) <- paste0('edgeR_offset_', names(HsDtuBenchmark_edgeR_offset))
names(HsDtuBenchmark_DESeq2_offset) <- paste0('DESeq2_offset_', names(HsDtuBenchmark_DESeq2_offset))

### Save result
save(HsDtuBenchmark_satuRn,
     HsDtuBenchmark_DEXSeq,
     HsDtuBenchmark_edgeRdiffsplice,
     HsDtuBenchmark_edgeR_offset,
     HsDtuBenchmark_DESeq2_offset,
     file="./Data/Hsapiens_DTU_results_count_new.Rdata")
```

```{r}
load("./Data/Hsapiens_DTU_results_count_new.Rdata")
```



